---
title: "Towards Biophysically-informed Drought Indicators Across US Ecosystems - Figures"
author: "Kesondra Key"
date: "2026-03-29"
output: 
  pdf_document: default
---

IMPORTANT: 
Change "E:/Github/diurnal-divergence-factor/data/" to your own file location in the setup

Then click "Knit" at the tob of the .Rmd file in RStudio

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

#Relevant packages installation
packages <- c(
  "dplyr", "tidyr", "ggplot2", "ggdist",
  "lubridate", "reshape2", "naniar", "BBmisc",
  "ggrepel", "gghighlight", "ggtext",
  "gridExtra", "multcomp", "multcompView",
  "readr", "stringr", "tinytex", "knitr"
)

to_install <- packages[!packages %in% installed.packages()[, "Package"]]

if (length(to_install) > 0) {
  install.packages(to_install)
}

invisible(lapply(packages, library, character.only = TRUE))

# Libraries
library(dplyr); library(tidyr); library(ggplot2); library(ggdist)
library(lubridate); library(reshape2); library(naniar); library(BBmisc)
library(ggrepel); library(gghighlight); library(ggtext)
library(gridExtra); library(grid); library(multcomp); library(multcompView)
library(readr); library(stringr); library(tinytex); library(knitr)

#IMPORTANT: Relevant File paths - change this to your own directory

# Dataset input dir
daily_path <- "E:/Github/diurnal-divergence-factor/data/Masterfile.csv"
SRSfile <- read.csv("E:/Github/diurnal-divergence-factor/data/US-SRS.csv")

# Figure output dir
output_dir <- "E:/Github/diurnal-divergence-factor/figures"

#Continue with setup
Daily <- readr::read_csv(daily_path, show_col_types = FALSE)

Daily <- Daily %>%
  mutate(
    Date = as.Date(Date),
    Drought = factor(
      dplyr::coalesce(Drought, ifelse(!is.na(Cluster) & Cluster > 0, "Drought", "ND")),
      levels = c("ND","Drought")
    ),
    drought_cat = factor(dplyr::coalesce(drought_cat, "ND"),
                         levels = c("ND","LS","MS")),
    AI_cat = factor(AI_cat,
                    levels = c("Humid","Semi-Arid","Arid"),
                    ordered = TRUE),
    Veg = factor(Veg, levels = c("DBF","ENF","SHR","GRA"))
  )

# Map p-values to stars
p_to_stars <- function(p) {
  if (is.na(p)) return("ns")
  if (p < 0.001) return("***")
  if (p < 0.01)  return("**")
  if (p < 0.05)  return("*")
  "ns"
}

# Tukey letters (for 3+ groups) per data.frame
tukey_letters <- function(dat, x, y) {
  dat <- dat[!is.na(dat[[x]]) & !is.na(dat[[y]]), , drop = FALSE]
  # group info
  g    <- factor(dat[[x]])
  levs <- levels(g)
  k    <- nlevels(g)

  make_tbl <- function(names_vec, labs_vec) {
    tibble::tibble(!!rlang::sym(x) := names_vec, .label = labs_vec)
  }

  if (k < 1) {
    return(make_tbl(character(0), character(0)))
  }
  if (k == 1) {
    return(make_tbl(levs, NA_character_))
  }
  if (k == 2) {
    p <- tryCatch(stats::t.test(dat[[y]] ~ g)$p.value, error = function(e) NA_real_)
    labs <- if (is.na(p) || p >= 0.05) c("a","a") else c("a","b")
    names(labs) <- levs
    return(make_tbl(names(labs), unname(labs)))
  }

  a  <- stats::aov(stats::as.formula(paste(y, "~", x)), data = dat)
  hs <- tryCatch(stats::TukeyHSD(a)[[1]][, "p adj"], error = function(e) NULL)

  if (is.null(hs)) {
    labs_full <- setNames(rep("a", k), levs)
  } else {
    tuk <- multcompView::multcompLetters(hs)$Letters   # named char vector
    labs_full <- setNames(rep("a", k), levs)
    common <- intersect(names(tuk), names(labs_full))
    labs_full[common] <- tuk[common]
  }

  make_tbl(names(labs_full), unname(labs_full))
}

theme_ddf_facet <- theme_classic() +
  theme(
    legend.position = "none",
    axis.line = element_line(linewidth = 0.8, color = "black"),
    axis.text.x = element_text(size = 16, face = "bold", color = "black"),
    axis.text.y = element_text(size = 16, face = "bold", color = "black"),
    axis.title.x = element_text(size = 16, face = "bold", color = "black"),
    axis.title.y = element_text(size = 16, face = "bold", color = "black"),
    axis.ticks = element_line(color = "black", linewidth = 0.8),
    axis.text = element_text(size = 14, face = "bold", color = "black"),
    axis.title = element_text(size = 13, face = "bold", color = "black"),
    strip.text = element_text(size = 12, face = "bold", color = "black"),
    panel.grid.major.x = element_line(linewidth = 0.4, color = "grey90"),
    panel.grid.minor.x = element_line(linewidth = 0.3, color = "grey95"),
    plot.title = element_text(size = 15, face = "bold", hjust = 0.5)
  )

```

## Figure 1. Relative ET VPD plot - Note: this plot was edited in powerpoint after this initial step
```{r Fig1, fig.cap= "Illustration of the DDF using observed eddy covariance data from a woody savanna site in Arizona (Ameriflux site: US-SRS). Blue depicts ET, with VPD in red, each normalized from 0 to 1 over the day (y-axis) and averaged by hour (x-axis). Panel A shows the typical diurnal sequence during a non-drought period (May-June 2016), with peak ET occurring during midday and peak VPD in the late afternoon. Panel B shows the same time of year during a drought (May-June 2013), where peak ET shifts earlier while VPD timing remains unchanged. This shift illustrates how drought advances peak ET timing relative to peak VPD and forms the conceptual basis for the DDF"}

knitr::opts_chunk$set(
  echo = FALSE,     
  warning = FALSE,   
  message = FALSE    
)
#US-SRS 
#Drought of interest:  May 5 to June 30 2013
#Non-Drought Year: 2016

#Variables of interest:
SRSfile$datetime <- SRSfile$DateTime
SRSfile$LE <- SRSfile$LE_f
SRSfile$NETRAD <- SRSfile$NETRAD
SRSfile$VPD_PI <- SRSfile$VPD_f

SRSDR <- SRSfile %>%
  dplyr::filter(datetime >= "2013-05-05 00:00:00" & datetime < "2013-06-30 00:00:00")%>%
  mutate(yday = yday(datetime)) %>%
  dplyr::select(datetime, Hour, yday, LE, NETRAD, VPD_PI) %>%
  replace_with_na_all(condition = ~.x == -9999) %>%
  dplyr::rename("ET" = LE) %>%
  dplyr::rename("Q" = NETRAD) %>%
  dplyr::rename("D" = VPD_PI) %>%
  dplyr::filter(Hour > 7 & Hour < 19) %>%
  group_by(Hour) %>%
  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
  mutate(ET = normalize(ET, method = "range", range = c(0, 1))) %>%
  mutate(Q = normalize(Q, method = "range", range = c(0, 1))) %>%
  mutate(D = normalize(D, method = "range", range = c(0, 1))) %>%
  dplyr::select(-Q) %>%
  melt(., id.vars = c("yday", "Hour")) %>%
  ggplot(aes(x = Hour, y = value, group = variable, colour = variable)) +
  geom_smooth(size = 2, span = 0.4, se = FALSE) +
  scale_color_manual(values = c("#42B2EB", "#EB7B42")) +
  xlab("Hour") +
  scale_x_continuous(breaks = c(8, 12, 16), labels = c("8 AM", "12 PM", "4 PM")) +
  scale_y_continuous(breaks = c(0, 1), labels = c("0", "1")) +
  theme_bw() +
  labs(y = "<span style='font-size:14pt;font-weight:bold;color:black'>Relative </span><span style='font-size:14pt;font-weight:bold;color:#42B2EB'>ET</span><span style='font-size:14pt;font-weight:bold;color:black'> and </span><span style='font-size:14pt;font-weight:bold;color:#EB7B42'>VPD</span>",
       y.subtitle = "Relative",
       title = "US-SRS LS Drought \n9/22/2019 to 10/5/2019") +
  theme(axis.title.y = element_markdown(),
        axis.text = element_text(size = 14, face = "bold"),
        axis.title.x = element_text(size = 14, face = "bold"),
        legend.position = "none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  geom_vline(data = . %>% filter(variable == "ET") %>% slice(which.max(value)),
             aes(xintercept = Hour), linetype = "dashed", color = "#42B2EB", size = 1) +
  geom_vline(data = . %>% filter(variable == "D") %>% slice(which.max(value)),
             aes(xintercept = Hour), linetype = "dashed", color = "#EB7B42", size = 1)

#Non-Drought Year of interest: 2016
SRSGR <- SRSfile %>%
  dplyr::filter(datetime >= "2016-05-05 00:00:00" & datetime < "2016-06-30 00:00:00")%>%
  mutate(yday = yday(datetime)) %>%
  dplyr::select(datetime, Hour, yday, LE, NETRAD, VPD_PI) %>%
  replace_with_na_all(condition = ~.x == -9999) %>%
  dplyr::rename("ET" = LE) %>%
  dplyr::rename("Q" = NETRAD) %>%
  dplyr::rename("D" = VPD_PI) %>%
  dplyr::filter(Hour > 7 & Hour < 19) %>%
  group_by(Hour) %>%
  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
  mutate(ET = normalize(ET, method = "range", range = c(0, 1))) %>%
  mutate(Q = normalize(Q, method = "range", range = c(0, 1))) %>%
  mutate(D = normalize(D, method = "range", range = c(0, 1))) %>%
  dplyr::select(-Q) %>%
  melt(., id.vars = c("yday", "Hour")) %>%
  ggplot(aes(x = Hour, y = value, group = variable, colour = variable)) +
  geom_smooth(size = 2, span = 0.4, se = FALSE) +
  scale_color_manual(values = c("#42B2EB", "#EB7B42")) +
  xlab("Hour") +
  scale_x_continuous(breaks = c(8, 12, 16), labels = c("8 AM", "12 PM", "4 PM")) +
  scale_y_continuous(breaks = c(0, 1), labels = c("0", "1")) +
  theme_bw() +
  labs(y = "<span style='font-size:14pt;font-weight:bold;color:black'>Relative </span><span style='font-size:14pt;font-weight:bold;color:#42B2EB'>ET</span><span style='font-size:14pt;font-weight:bold;color:black'> and </span><span style='font-size:14pt;font-weight:bold;color:#EB7B42'>VPD</span>",
       y.subtitle = "Relative",
       title = "US-MOZ ND Year\n9/22/2010 to 10/5/2010") +
  theme(axis.title.y = element_markdown(),
        axis.text = element_text(size = 14, face = "bold"),
        axis.title.x = element_text(size = 14, face = "bold"),
        legend.position = "none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  geom_vline(data = . %>% filter(variable == "ET") %>% slice(which.max(value)),
             aes(xintercept = Hour), linetype = "dashed", color = "#42B2EB", size = 1) +
  geom_vline(data = . %>% filter(variable == "D") %>% slice(which.max(value)),
             aes(xintercept = Hour), linetype = "dashed", color = "#EB7B42", size = 1)

plot1 <- SRSDR #Drought
plot2 <- SRSGR #Non-Drought

# Create titles for each plot
title1 <- "Drought Period\n5/5/2013 to 6/30/2013"
title2 <- "Non-Drought Year\n5/5/2016 to 6/30/2016"
overall_title <- "Site: US-SRS"

title_theme <- theme(plot.title = element_text(face = "bold", hjust = 0.5))

# Arrange the plots side by side
grid.arrange(
  plot2 + ggtitle(title2) + title_theme, 
  plot1 + ggtitle(title1) + title_theme,
  nrow = 1,
  top = textGrob(overall_title, gp = gpar(fontsize = 16, fontface = "bold"))
)

#Save
output_path <- file.path(output_dir, "Figure_1.png")

# Sava as a PNG 
ggsave(output_path, width = 7, height = 4, dpi = 300, bg = "white", units = "in", 
       plot = grid.arrange(
         plot2 + ggtitle(title2) + title_theme, 
         plot1 + ggtitle(title1) + title_theme,
         nrow = 1,
         top = textGrob(overall_title, gp = gpar(fontsize = 16, fontface = "bold"))
       ))
```

### Note:
Figure 2 code is in the "Figure_2_REQUIRES_EXTERNAL_FILE.R" script, which requires downloading the aridity index .tif from the original source. 
-> Information and citation in the script mentioned above. 

\newpage
## Figure 3. Drought versus Non-Drought
```{r Fig3, fig.cap= "Comparison of DDF distributions across drought vs. non-drought conditions for 40 pooled flux tower sites. Dotted horizontal lines indicate mean ND timing. Boxplots show medians (black horizontal lines) and means (white points) while kernel density overlays  (bandwidth = 0.4) further describe the distributions. Statistical significance: A) Welch’s t-test (*** p < 0.001)"}

knitr::opts_chunk$set(
  echo = FALSE,     
  warning = FALSE,  
  message = FALSE   
)

# Drought vs Non-Drought
df <- Daily %>% filter(!is.na(Drought), !is.na(DDF)) %>%
  mutate(Drought = factor(Drought, levels = c("ND","Drought"))) %>%
  mutate(
    x_pos = as.numeric(Drought) + dplyr::case_when(
      Drought == "ND"      ~ -0.2,
      Drought == "Drought" ~  0.0
    )
  )

bw_use <- 0.40
dens   <- ggdist::density_unbounded(bandwidth = bw_use)

nd_line <- df %>%
  filter(Drought == "ND") %>%
  summarise(nd_med = median(DDF, na.rm = TRUE))

# t-test + stars
ttest <- tryCatch(stats::t.test(DDF ~ Drought, data = df),
                  error = function(e) list(p.value = NA_real_))

sig_label <- dplyr::case_when(
  ttest$p.value < 0.001 ~ "***",
  ttest$p.value < 0.01  ~ "**",
  ttest$p.value < 0.05  ~ "*",
  TRUE                  ~ "ns"
)

y_max <- max(df$DDF, na.rm = TRUE)
span  <- diff(range(df$DDF, na.rm = TRUE))
y_star <- y_max + 0.08 * span
x_star <- mean(c(0.8, 2.0))  

droughtplot <- ggplot(df, aes(x = x_pos, y = DDF, fill = Drought, group = Drought)) +
  ggplot2::geom_hline(
    data = nd_line,
    aes(yintercept = nd_med),
    linetype = "dashed",
    color = "gray30",
    linewidth = 0.8
  ) +

  ggdist::stat_halfeye(
    data = ~subset(.x, Drought == "ND"),
    density = dens, .width = 0, alpha = 0.5,
    slab_color = "#17BEBB", slab_linewidth = 1.1
  ) +
  ggdist::stat_halfeye(
    data = ~subset(.x, Drought == "Drought"),
    density = dens, .width = 0, alpha = 0.5,
    slab_color = "#E4572E", slab_linewidth = 1.1
  ) +

  geom_boxplot(width = 0.35, linewidth = 0.8, color = "black", outlier.shape = NA) +
  stat_summary(fun = mean, geom = "point",
               shape = 21, size = 2, color = "black", fill = "white", stroke = 0.8) +
  scale_y_continuous(
    breaks = seq(-10, 10, by = 2)
  ) +

  coord_flip(clip = "off") +
  theme_ddf_facet +
  scale_x_continuous(
    breaks = c(0.8, 2.0),
    labels = c("ND","Drought")
  ) +

  scale_fill_manual(values = c("ND" = "#17BEBB", "Drought" = "#E4572E")) +
  labs(
    y = "DDF (hours)", 
    x = "",
  ) +
  annotate("text", x = x_star, y = y_star, label = sig_label,
           color = "black", size = 7, fontface = "bold")

print(droughtplot)

ggsave(file.path(output_dir, "Figure_3.png"),
       droughtplot, width = 8, height = 2.5, dpi = 300, bg = "white", units = "in")
```

\newpage
## Figure 4. Drought Severity
```{r Fig4, fig.cap = "Comparison of DDF distributions across drought severity categories, for 40 pooled flux tower sites. Dotted horizontal lines indicate mean ND timing. Boxplots show medians (black horizontal lines) and means (white points); density overlays Gaussian kernel estimates (bandwidth = 0.4). Statistical significance: Tukey HSD post-hoc comparisons (letters denote statistically significant groups). Outliers were removed for plotting purposes but not analyses"}
knitr::opts_chunk$set(
  echo = FALSE,      
  warning = FALSE,   
  message = FALSE   
)
Daily$drought_cat <- factor(Daily$drought_cat, levels = c("ND","LS","MS"))
dat3 <- Daily %>% filter(!is.na(drought_cat), !is.na(DDF))

bw_use <- 0.40
dens   <- ggdist::density_unbounded(bandwidth = bw_use)

dat3 <- dat3 %>%
  mutate(
    x_pos = as.numeric(drought_cat) + dplyr::case_when(
      drought_cat == "ND" ~ 0.0,
      drought_cat == "LS" ~ 0.2,
      drought_cat == "MS" ~ 0.4
    )
  )

nd_line <- dat3 %>%
  filter(drought_cat == "ND") %>%
  summarise(nd_med = median(DDF, na.rm = TRUE))

# Tukey letters + label positions (aligned to x_pos)
letters_df <- tukey_letters(dat3, x = "drought_cat", y = "DDF")

y_tops <- dat3 %>%
  group_by(drought_cat) %>%
  summarise(y_top = max(DDF, na.rm = TRUE), .groups = "drop")

span <- diff(range(dat3$DDF, na.rm = TRUE))

lvl <- c("ND","LS","MS")
offs <- c("ND" = 0.0, "LS" = 0.2, "MS" = 0.4)

lvl <- c("ND","LS","MS")
offs <- c("ND" = 0.0, "LS" = 0.2, "MS" = 0.4)

letters_plot <- y_tops %>%
  mutate(drought_cat = as.character(drought_cat)) %>%
  left_join(
    letters_df %>% mutate(drought_cat = as.character(drought_cat)),
    by = "drought_cat"
  ) %>%
  mutate(
    idx   = as.numeric(factor(drought_cat, levels = lvl)),
    x_pos = idx + offs[drought_cat],
    y_lab = y_top + 0.06 * span
  )

droughtcats <- ggplot(dat3, aes(x = x_pos, y = DDF, fill = drought_cat, group = drought_cat)) +
  ggplot2::geom_hline(
    data = nd_line,
    aes(yintercept = nd_med),
    linetype = "dashed",
    color = "gray30",
    linewidth = 0.8
  ) +

  ggdist::stat_halfeye(
    data = ~subset(.x, drought_cat == "ND"),
    density = dens, .width = 0, alpha = 0.5,
    slab_color = "#56B4E9", slab_linewidth = 1.1
  ) +
  ggdist::stat_halfeye(
    data = ~subset(.x, drought_cat == "LS"),
    density = dens, .width = 0, alpha = 0.5,
    slab_color = "#E69F00", slab_linewidth = 1.1
  ) +
  ggdist::stat_halfeye(
    data = ~subset(.x, drought_cat == "MS"),
    density = dens, .width = 0, alpha = 0.5,
    slab_color = "#D55E00", slab_linewidth = 1.1
  ) +

  geom_boxplot(width = 0.35, linewidth = 0.8, color = "black", outlier.shape = NA) +

  stat_summary(fun = mean, geom = "point",
               shape = 21, size = 2, color = "black", fill = "white", stroke = 0.8) +

  scale_y_continuous(
    breaks = seq(-10, 10, by = 2)
  ) +
  coord_flip(clip = "off") +
  theme_ddf_facet +

  scale_x_continuous(
    breaks = c(1.0, 2.2, 3.4),
    labels = c("ND","LS","MS")
  ) +
  scale_fill_manual(values = c("ND"="#56B4E9","LS"="#E69F00","MS"="#D55E00")) +
  labs(
    y = "Daily DDF (hours)", x = "",
  ) +
  

  geom_text(
    data = letters_plot,
    inherit.aes = FALSE,
    aes(x = x_pos, y = y_lab, label = .label),
    fontface = "bold", size = 5, color = "black"
  )

print(droughtcats)

ggsave(file.path(output_dir,"Figure_4.png"),
       droughtcats, width=8, height=3.5, dpi=300, bg="white", units="in")

```

\newpage
## Figure 5. Aridity - Drought vs. non-drought
```{r Fig5,  fig.cap= "Comparison of DDF distributions across aridity classes during drought and non-drought conditions. Dotted horizontal lines indicate mean ND timing. Boxplots show medians (black horizontal lines) and means (white points); density overlays are Gaussian kernel estimates (bandwidth = 0.4). Asterisks indicate significant differences between drought and ND within each aridity class (Welch’s t-test, *** p < 0.001). Outliers were removed for plotting purposes but not analyses"}
knitr::opts_chunk$set(
  echo = FALSE,      
  warning = FALSE,   
  message = FALSE   
)

AI_cat_df <- Daily %>%
  filter(!is.na(AI_cat), !is.na(Drought), !is.na(DDF)) %>%
  mutate(
    Drought = factor(Drought, levels = c("ND","Drought")),
    AI_cat  = factor(AI_cat, levels = c("Arid", "Semi-Arid", "Humid")),
    x_pos   = as.numeric(Drought) + dplyr::case_when(
      Drought == "ND"      ~ -0.2,
      Drought == "Drought" ~  0.0
    )
  )

bw_use <- 0.4
dens   <- ggdist::density_unbounded(bandwidth = bw_use)

sig_AI_cat <- AI_cat_df %>%
  group_by(AI_cat) %>%
  summarise(
    p = tryCatch(t.test(DDF ~ Drought)$p.value, error = function(e) NA_real_),
    y_max = max(DDF, na.rm = TRUE),
    span  = diff(range(DDF, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  mutate(
    star  = vapply(p, p_to_stars, character(1)),
    y_bar = y_max + 0.08 * span,
    x_bar = mean(c(0.8, 2.0))  
  )

nd_line_ai <- AI_cat_df %>%
  filter(Drought == "ND") %>%
  group_by(AI_cat) %>%
  summarise(nd_med = median(DDF, na.rm = TRUE), .groups = "drop")

AI_catDroughtPlot <- ggplot(AI_cat_df, aes(x = x_pos, y = DDF, fill = Drought, group = Drought)) +
  ggplot2::geom_hline(
    data = nd_line_ai,
    aes(yintercept = nd_med),
    linetype = "dashed",
    color = "gray30",
    linewidth = 0.8
  ) +

  ggdist::stat_halfeye(
    data = ~subset(.x, Drought == "ND"),
    density = dens, .width = 0, alpha = 0.5,
    slab_color = "#17BEBB", slab_linewidth = 1.1
  ) +
  ggdist::stat_halfeye(
    data = ~subset(.x, Drought == "Drought"),
    density = dens, .width = 0, alpha = 0.5,
    slab_color = "#E4572E", slab_linewidth = 1.1
  ) +

  geom_boxplot(width = 0.35, linewidth = 0.8, color = "black", outlier.shape = NA) +
  stat_summary(fun = mean, geom = "point",
               shape = 21, size = 2, color = "black", fill = "white", stroke = 0.8) +
    scale_y_continuous(
    breaks = seq(-10, 10, by = 2)
    )+

  coord_flip(clip = "off") +
  theme_ddf_facet +

  scale_x_continuous(
    breaks = c(0.8, 2.0),
    labels = c("ND","Drought")
  ) +

  scale_fill_manual(values = c("ND" = "#17BEBB", "Drought" = "#E4572E")) +
  facet_wrap(~AI_cat, nrow = 3, ncol = 1) +

  labs(
    y = "Daily DDF (hours)", x = "",
  ) +

  geom_text(
    data = sig_AI_cat,
    inherit.aes = FALSE,
    aes(x = x_bar, y = y_bar, label = star),
    fontface = "bold", size = 7
  )

print(AI_catDroughtPlot)

ggsave(file.path(output_dir,"Figure_5.png"),
       AI_catDroughtPlot, width=8, height=6, dpi=300, bg="white", units="in")

```

\newpage
## Figure 6. Delta DDF barplots organized by aridity
```{r Fig6, fig.cap= "Site-level drought effects on mean DDF (delta DDF  = DDFdrought – DDFND), organized by site aridity class then by magnitude of the DDF signal. Negative values indicate earlier ET peak timing during drought"}

knitr::opts_chunk$set(
  echo = FALSE,     
  warning = FALSE,   
  message = FALSE    
)

library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpattern)

theme_ddf_facet <- theme_classic() +
  theme(
    legend.position = "none",
    axis.line = element_line(linewidth = 0.8, color = "black"),
    axis.text.x = element_text(size = 14, face = "bold", color = "black"),
    axis.text.y = element_text(size = 14, face = "bold", color = "black"),
    axis.ticks = element_line(color = "black", linewidth = 0.8),
    axis.title.y = element_text(size = 14, face = "bold", color = "black"),
    axis.title.x = element_text(size = 12, face = "bold", color = "black"),
    strip.text = element_text(size = 12, face = "bold", color = "black"),
    panel.grid.major.x = element_line(linewidth = 0.4, color = "grey90"),
    panel.grid.minor.x = element_line(linewidth = 0.3, color = "grey95"),
    plot.title = element_text(size = 15, face = "bold", hjust = 0.5)
  )

# 1) Prep
df <- Daily %>%
  filter(!is.na(Drought), !is.na(DDF), !is.na(AI_cat), !is.na(Veg), !is.na(mean_AI)) %>%
  mutate(
    Drought = factor(Drought, levels = c("ND", "Drought")),
    Veg     = factor(Veg, levels = c("DBF", "ENF", "GRA", "SHR")),
    AI_cat  = factor(AI_cat, levels = c("Arid", "Humid"))
  )

# 2) One veg / AI_cat / mean_AI per site
site_veg <- df %>%
  count(sitename, Veg, name = "n") %>%
  arrange(sitename, desc(n)) %>%
  group_by(sitename) %>%
  slice(1) %>%
  ungroup() %>%
  dplyr::select(sitename, Veg)

site_ai <- df %>%
  count(sitename, AI_cat, name = "n") %>%
  arrange(sitename, desc(n)) %>%
  group_by(sitename) %>%
  slice(1) %>%
  ungroup() %>%
  dplyr::select(sitename, AI_cat)

site_mean_ai <- df %>%
  group_by(sitename) %>%
  summarise(mean_AI = mean(mean_AI, na.rm = TRUE), .groups = "drop")

# 3) Mean DDF per site × class, then delta
delta_site <- df %>%
  group_by(sitename, Drought) %>%
  summarise(
    mean_DDF = mean(DDF, na.rm = TRUE),
    n_days   = n(),
    .groups  = "drop"
  ) %>%
  pivot_wider(
    names_from = Drought,
    values_from = c(mean_DDF, n_days)
  ) %>%
  mutate(delta_DDF = mean_DDF_Drought - mean_DDF_ND) %>%
  left_join(site_veg, by = "sitename") %>%
  left_join(site_ai, by = "sitename") %>%
  left_join(site_mean_ai, by = "sitename") %>%
  filter(!is.na(delta_DDF), !is.na(Veg), !is.na(AI_cat), !is.na(mean_AI)) %>%
  arrange(delta_DDF) %>%
  mutate(
    site_order = row_number(),
    sitename = factor(sitename, levels = sitename)
  )

ai_cols <- c(
  "Arid" = "#E4572E",
  "Humid" = "#17BEBB"
)

p1 <- ggplot(
  delta_site,
  aes(x = site_order, y = delta_DDF, fill = AI_cat, pattern = AI_cat)
) +
  geom_hline(
    yintercept = c(-0.5, 0.5),
    linewidth = 0.6,
    color = "gray50",
    linetype = "dashed"
  ) +
  geom_hline(yintercept = 0, linewidth = 0.8, color = "gray30") +
 geom_col(width = 0.85, color = "black") +
  scale_fill_manual(values = ai_cols) +
  scale_pattern_manual(values = c("Arid" = "none", "Humid" = "stripe")) +
  scale_x_continuous(
    breaks = delta_site$site_order,
    labels = delta_site$sitename,
    expand = expansion(mult = c(0.01, 0.03))
  ) +
  scale_y_continuous(
    breaks = seq(-1, 1, 0.5)
  ) +
  theme_ddf_facet +
  theme(
    axis.line = element_line(linewidth = 0.8, color = "black"),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10, color = "black"),
    axis.text.y = element_text(size = 10, face = "bold", color = "black"),
    axis.title = element_text(size = 12, face = "bold"),
    legend.position = "bottom",
    legend.text = element_text(face = "bold", color = "black", size = 11),
    legend.title = element_blank(),
    plot.margin = margin(20, 10, 10, 10)
  ) +
  labs(
    x = NULL,
    y = "ΔDDF (hours)"
  )

print(p1)

ggsave(
  file.path(output_dir, "Figure_6.png"),
  p1,
  width = 12,
  height = 8,
  dpi = 600,
  bg = "white",
  units = "in",
  device = ragg::agg_png
)

```

\newpage
## Figure 7. Aridity - drought severity
```{r Fig7, fig.cap= "Comparison of DDF distributions across drought severity categories (ND LS, MS) within each aridity class. Dotted horizontal lines indicate mean ND timing for each aridity class. Boxplots show medians (black horizontal lines) and means (white points); density overlays are Gaussian kernel estimates (bandwidth = 0.4). Letters indicate significantly different groups based on Tukey HSD post-hoc tests (p < 0.05). Arid sites had no MS droughts detected. Outliers were removed for plotting purposes only"}

bw_use <- 0.4
dens   <- ggdist::density_unbounded(bandwidth = bw_use)

v3 <- Daily %>%
  filter(!is.na(AI_cat), !is.na(drought_cat), !is.na(DDF)) %>%
  mutate(
    drought_cat = factor(drought_cat, levels = c("ND","LS","MS")),
    AI_cat      = factor(AI_cat, levels = c("Arid", "Semi-Arid", "Humid")),
    x_pos = as.numeric(drought_cat) + dplyr::case_when(
      drought_cat == "ND" ~ 0.0,
      drought_cat == "LS" ~ 0.2,
      drought_cat == "MS" ~ 0.4
    )
  )

nd_line_ai <- v3 %>%
  filter(drought_cat == "ND") %>%
  group_by(AI_cat) %>%
  summarise(nd_med = median(DDF, na.rm = TRUE), .groups = "drop")

letters_by_AI_cat <- v3 %>%
  group_by(AI_cat) %>%
  group_modify(~{
    labs <- tukey_letters(.x, x = "drought_cat", y = "DDF")
    y_top <- .x %>%
      group_by(drought_cat) %>%
      summarise(y_top = max(DDF, na.rm = TRUE), .groups = "drop")
    
    span <- diff(range(.x$DDF, na.rm = TRUE))
    
    dplyr::left_join(labs, y_top, by = "drought_cat") %>%
      mutate(
        y_lab = y_top + 0.06 * span,
        x_pos = as.numeric(factor(drought_cat, levels = c("ND","LS","MS"))) +
          dplyr::case_when(
            drought_cat == "ND" ~ 0.0,
            drought_cat == "LS" ~ 0.2,
            drought_cat == "MS" ~ 0.4
          )
      )
  }) %>%
  ungroup()

AI_catSevCats <- ggplot(v3, aes(x = x_pos, y = DDF, fill = drought_cat, group = drought_cat)) +
  ggplot2::geom_hline(
    data = nd_line_ai,
    aes(yintercept = nd_med),
    linetype = "dashed",
    color = "gray30",
    linewidth = 0.8
  ) +
  
  ggdist::stat_halfeye(
    data = ~subset(.x, drought_cat == "ND"),
    density = dens, .width = 0, alpha = 0.5,
    slab_color = "#56B4E9", slab_linewidth = 1.1
  ) +
  ggdist::stat_halfeye(
    data = ~subset(.x, drought_cat == "LS"),
    density = dens, .width = 0, alpha = 0.5,
    slab_color = "#E69F00", slab_linewidth = 1.1
  ) +
  ggdist::stat_halfeye(
    data = ~subset(.x, drought_cat == "MS"),
    density = dens, .width = 0, alpha = 0.5,
    slab_color = "#D55E00", slab_linewidth = 1.1
  ) +
  
  geom_boxplot(width = 0.35, linewidth = 0.8, color = "black", outlier.shape = NA) +
  stat_summary(fun = mean, geom = "point",
               shape = 21, size = 2, color = "black", fill = "white", stroke = 0.8) +
  scale_y_continuous(
    breaks = seq(-10, 10, by = 2)
    )+
  
  coord_flip(clip = "off") +
  theme_ddf_facet +
  
  scale_x_continuous(
    breaks = c(1.0, 2.2, 3.4),
    labels = c("ND","LS","MS")
  ) +
  
  scale_fill_manual(values = c("ND"="#56B4E9","LS"="#E69F00","MS"="#D55E00")) +
  facet_wrap(~AI_cat, nrow = 3, ncol = 1) +
  
  labs(
    y = "Daily DDF (hours)", x = "",
  ) +
  
  geom_text(
    data = letters_by_AI_cat,
    aes(x = x_pos, y = y_lab, label = .label),
    inherit.aes = FALSE,
    fontface = "bold", size = 5
  )

print(AI_catSevCats)

ggsave(file.path(output_dir,"Figure_7.png"),
       AI_catSevCats, width=8, height=8, dpi=300, bg="white", units="in")

```

\newpage
# SUPPLEMENTARY PLOTS

## Figure S1. Aridity category comparison
```{r FigS1, fig.cap= "Comparison of DDF distributions across aridity gradients. Boxplots show medians (black horizontal lines) and means (white points); density overlays Gaussian kernel estimates (bandwidth = 0.4). Statistical significance: Welch’s t-test (*** p <0.001). Outliers were removed for plotting purposes but not analyses"}
pal_ai <- c(
  "Arid" = "#E4572E",
  "Humid" = "#17BEBB"
)

bw_use <- 0.4
dens   <- ggdist::density_unbounded(bandwidth = bw_use)

lvl  <- c("Humid","Arid")
offs <- c(Humid = 0.0, Arid = 0.4)

dat3 <- Daily %>%
  filter(!is.na(AI_cat), !is.na(DDF)) %>%
  mutate(
    AI_cat = factor(AI_cat, levels = lvl),
    idx   = as.numeric(AI_cat),
    x_pos = idx + offs[as.character(AI_cat)]
  )

# one overall Arid vs Humid test
p_ai <- tryCatch(
  t.test(DDF ~ AI_cat, data = dat3)$p.value,
  error = function(e) NA_real_
)

sig_AI_cat <- dat3 %>%
  summarise(
    y_max = max(DDF, na.rm = TRUE),
    y_min = min(DDF, na.rm = TRUE)
  ) %>%
  mutate(
    span = y_max - y_min,
    star = p_to_stars(p_ai),
    y_bar = y_max + 0.08 * span,
    x_bar = mean(c(1.0, 2.4))
  )

ai_plot <- ggplot(dat3, aes(x = x_pos, y = DDF, fill = AI_cat, group = AI_cat)) +
  
  ggdist::stat_halfeye(
    data = ~subset(.x, AI_cat == "Humid"),
    density = dens, .width = 0, alpha = 0.35,   
    slab_color = pal_ai["Humid"], slab_linewidth = 1.1
  ) +
  
  ggdist::stat_halfeye(
    data = ~subset(.x, AI_cat == "Arid"),
    density = dens, .width = 0, alpha = 0.35,
    slab_color = pal_ai["Arid"], slab_linewidth = 1.1
  ) +
  
  geom_boxplot(
    width = 0.35, linewidth = 0.8, color = "black", outlier.shape = NA,
    alpha = 0.90                              
  ) +
  
  stat_summary(
    fun = mean, geom = "point",
    shape = 21, size = 2.2,
    color = "black", fill = "white",
    stroke = 1.1                               
  ) +
  
  scale_y_continuous(breaks = seq(-10, 10, by = 2)) +
  coord_flip(clip = "off") +
  theme_ddf_facet +
  
  scale_x_continuous(
    breaks = c(1.0, 2.2),
    labels = c("Humid","Arid")       
  ) +
  
  scale_fill_manual(
    values = pal_ai,
    breaks = lvl,
    labels = c(Humid="Humid", Arid="Arid") 
  ) +
  
  labs(
    y = "Daily DDF (hours)", x = "",
  ) +
  
    geom_text(
    data = sig_AI_cat,
    inherit.aes = FALSE,
    aes(x = x_bar, y = y_bar, label = star),
    fontface = "bold", size = 5
  )

print(ai_plot)

ggsave(
  file.path(output_dir, "Figure_S1.png"),
  ai_plot, width = 8, height = 3.5, dpi = 300, bg = "white", units = "in"
)

```

## Figure S2. Vegetation plots
```{r FigS2, fig.cap= "Comparison of DDF distributions across vegetation gradients. Boxplots show medians (black horizontal lines) and means (white points); density overlays Gaussian kernel estimates (bandwidth = 0.4). Statistical significance: Tukey HSD post-hoc comparisons (letters denote statistically significant groups). Outliers were removed for plotting purposes but not analyses"}
pal_ai <- c(
  "DBF" = "#753bbd",
  "ENF" = "#2F80ED",
  "GRA" = "#2dc84d", 
  "SHR" = "#ff7f41"
)

bw_use <- 0.4
dens   <- ggdist::density_unbounded(bandwidth = bw_use)

lvl = c("DBF", "ENF", "GRA", "SHR")
offs <- c(DBF = 0.0, ENF = 0.2, GRA = 0.4, SHR = 0.6)

dat3 <- Daily %>%
  filter(!is.na(Veg), !is.na(DDF)) %>%
  mutate(
    Veg = factor(Veg, levels = lvl),
    idx   = as.numeric(Veg),
    x_pos = idx + offs[as.character(Veg)]
  )

letters_df <- tukey_letters(dat3, x = "Veg", y = "DDF")

y_tops <- dat3 %>%
  group_by(Veg) %>%
  summarise(y_top = max(DDF, na.rm = TRUE), .groups = "drop")

span <- diff(range(dat3$DDF, na.rm = TRUE))

letters_plot <- y_tops %>%
  mutate(Veg = as.character(Veg)) %>%
  left_join(letters_df %>% mutate(Veg = as.character(Veg)), by = "Veg") %>%
  mutate(
    idx   = as.numeric(factor(Veg, levels = lvl)),
    x_pos = idx + offs[Veg],
    y_lab = y_top + 0.06 * span
  )

veg_plot <- ggplot(dat3, aes(x = x_pos, y = DDF, fill = Veg, group = Veg)) +
  ggdist::stat_halfeye(
    data = ~subset(.x, Veg == "DBF"),
    density = dens, .width = 0, alpha = 0.35,  
    slab_color = pal_ai["DBF"], slab_linewidth = 1.1
  ) +
  ggdist::stat_halfeye(
    data = ~subset(.x, Veg == "ENF"),
    density = dens, .width = 0, alpha = 0.35,
    slab_color = pal_ai["ENF"], slab_linewidth = 1.1
  ) +

  ggdist::stat_halfeye(
    data = ~subset(.x, Veg == "GRA"),
    density = dens, .width = 0, alpha = 0.35, 
    slab_color = pal_ai["GRA"], slab_linewidth = 1.1
  ) +
  ggdist::stat_halfeye(
    data = ~subset(.x, Veg == "SHR"),
    density = dens, .width = 0, alpha = 0.35,
    slab_color = pal_ai["SHR"], slab_linewidth = 1.1
  ) +

  
  geom_boxplot(
    width = 0.35, linewidth = 0.8, color = "black", outlier.shape = NA,
    alpha = 0.90                                 
  ) +
  
  stat_summary(
    fun = mean, geom = "point",
    shape = 21, size = 2.2,
    color = "black", fill = "white",
    stroke = 1.1                                 
  ) +
  
  scale_y_continuous(breaks = seq(-10, 10, by = 2)) +
  coord_flip(clip = "off") +
  theme_ddf_facet +
  
  scale_x_continuous(
    breaks = c(1.0, 2.2, 3.4,4.6),
    labels = c("DBF", "ENF", "GRA", "SHR")   
  ) +
  
  scale_fill_manual(
    values = pal_ai,
    breaks = lvl
  ) +
  
  labs(
    y = "Daily DDF (hours)", x = "",
  ) +
  
  geom_text(
    data = letters_plot,
    inherit.aes = FALSE,
    aes(x = x_pos, y = y_lab, label = .label),
    fontface = "bold", size = 5
  )

print(veg_plot)

ggsave(
  file.path(output_dir, "Figure_S2.png"),
  veg_plot, width = 8, height = 6, dpi = 300, bg = "white", units = "in"
)

```

\newpage
## Figure S3. Vegetation drought vs non-drought plots
```{r FigS3,  fig.cap= "Comparison of DDF distributions and means during drought and non-drought (ND) conditions across Vegetation types. The dotted horizontal line indicates where the timing of peak ET and VPD occur during ND periods for each category on average. Boxplots show medians (black horizontal lines) and means (white points). Half-eye density overlays are unbounded Gaussian kernel density estimates (bandwidth = 0.4) to provide light smoothing for visualization. Significant differences of mean values (p < 0.05*, < 0.01**, < 0.001***) based on Welch’s t-test between groups are indicated by asterisks to the right of the boxplots. Outliers were removed for plotting purposes only"}
knitr::opts_chunk$set(
  echo = FALSE,      
  warning = FALSE,   
  message = FALSE   
)

bw_use <- 0.40
dens   <- ggdist::density_unbounded(bandwidth = bw_use)

veg_df <- Daily %>%
  filter(!is.na(Veg), !is.na(Drought), !is.na(DDF)) %>%
  mutate(
    Drought = factor(Drought, levels = c("ND","Drought")),
    Veg     = factor(Veg, levels = c("DBF","ENF","GRA","SHR")),
    x_pos   = as.numeric(Drought) + dplyr::case_when(
      Drought == "ND"      ~ -0.2,
      Drought == "Drought" ~  0.0
    )
  )

sig_veg <- veg_df %>%
  group_by(Veg) %>%
  summarise(
    p = tryCatch(t.test(DDF ~ Drought)$p.value, error = function(e) NA_real_),
    y_max = max(DDF, na.rm = TRUE),
    span  = diff(range(DDF, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  mutate(
    star  = vapply(p, p_to_stars, character(1)),
    y_bar = y_max + 0.08 * span,
    x_bar = mean(c(0.8, 2.0)) 
  )

nd_line_veg <- veg_df %>%
  filter(Drought == "ND") %>%
  group_by(Veg) %>%
  summarise(nd_med = median(DDF, na.rm = TRUE), .groups = "drop")

VegDroughtPlot <- ggplot(veg_df, aes(x = x_pos, y = DDF, fill = Drought, group = Drought)) +
  ggplot2::geom_hline(
    data = nd_line_veg,
    aes(yintercept = nd_med),
    linetype = "dashed",
    color = "gray30",
    linewidth = 0.8
  ) +

  ggdist::stat_halfeye(
    data = ~subset(.x, Drought == "ND"),
    density = dens, .width = 0, alpha = 0.5,
    slab_color = "#17BEBB", slab_linewidth = 1.1
  ) +
  ggdist::stat_halfeye(
    data = ~subset(.x, Drought == "Drought"),
    density = dens, .width = 0, alpha = 0.5,
    slab_color = "#E4572E", slab_linewidth = 1.1
  ) +

  geom_boxplot(width = 0.35, linewidth = 0.8, color = "black", outlier.shape = NA) +
  stat_summary(fun = mean, geom = "point",
               shape = 21, size = 2, color = "black", fill = "white", stroke = 0.8) +

  scale_y_continuous(
  breaks = seq(-10, 10, by = 2)
)+
  coord_flip(clip = "off") +
  theme_ddf_facet +

  scale_x_continuous(
    breaks = c(0.8, 2.0),
    labels = c("ND","Drought")
  ) +

  facet_wrap(~Veg, ncol = 2, nrow = 3) +
  scale_fill_manual(values = c("ND" = "#17BEBB", "Drought" = "#E4572E")) +

  labs(
    y = "Daily DDF (hours)", x = "",
  ) +

  geom_text(
    data = sig_veg,
    inherit.aes = FALSE,
    aes(x = x_bar, y = y_bar, label = star),
    fontface = "bold", size = 5
  )

print(VegDroughtPlot)

ggsave(file.path(output_dir,"Figure_S3.png"),
       VegDroughtPlot, width=10, height=6, dpi=300, bg="white", units="in")

```

## Figure S4. Vegetation drought severity plots
```{r FigS4, fig.cap= "Comparison of DDF distributions and means during each drought severity category across vegetation types. The dotted horizontal line indicates where the timing of peak ET and VPD occur during ND periods for each category on average. Boxplots show medians (black horizontal lines) and means (white points). Half-eye density overlays are unbounded Gaussian kernel density estimates (bandwidth = 0.4) to provide light smoothing for visualization. Letters denote differences in Tukey’s post hoc test to the right of the boxplots. Outliers were removed for plotting purposes only. MS sample sizes were small for several vegetation types (ENF: 2 sites; SHR: 3 sites); results should be interpreted with caution"}

v3 <- Daily %>%
  filter(!is.na(Veg), !is.na(drought_cat), !is.na(DDF)) %>%
  mutate(
    drought_cat = factor(drought_cat, levels = c("ND","LS","MS")),
    Veg = factor(Veg, levels = c("DBF","ENF","GRA","SHR")),
    x_pos = as.numeric(drought_cat) + dplyr::case_when(
      drought_cat == "ND" ~ 0.0,
      drought_cat == "LS" ~ 0.2,
      drought_cat == "MS" ~ 0.4
    )
  )

nd_line_vegsev <- v3 %>%
  filter(drought_cat == "ND") %>%
  group_by(Veg) %>%
  summarise(nd_med = median(DDF, na.rm = TRUE), .groups = "drop")

# Tukey letters + positions per Veg facet 
letters_by_veg <- v3 %>%
  group_by(Veg) %>%
  group_modify(~{
    labs <- tukey_letters(.x, x = "drought_cat", y = "DDF")
    y_top <- .x %>%
      group_by(drought_cat) %>%
      summarise(y_top = max(DDF, na.rm = TRUE), .groups = "drop")

    span <- diff(range(.x$DDF, na.rm = TRUE))

    dplyr::left_join(labs, y_top, by = "drought_cat") %>%
      mutate(
        y_lab = y_top + 0.06 * span,
        x_pos = as.numeric(factor(drought_cat, levels = c("ND","LS","MS"))) +
          dplyr::case_when(
            drought_cat == "ND" ~ 0.0,
            drought_cat == "LS" ~ 0.2,
            drought_cat == "MS" ~ 0.4
          )
      )
  }) %>%
  ungroup()

VegSevCats <- ggplot(v3, aes(x = x_pos, y = DDF, fill = drought_cat, group = drought_cat)) +
  
  ggplot2::geom_hline(
    data = nd_line_vegsev,
    aes(yintercept = nd_med),
    linetype = "dashed",
    color = "gray30",
    linewidth = 0.8
  ) +

  # Half-eyes (shared KDE)
  ggdist::stat_halfeye(
    data = ~subset(.x, drought_cat == "ND"),
    density = dens, .width = 0, alpha = 0.5,
    slab_color = "#56B4E9", slab_linewidth = 1.1
  ) +
  ggdist::stat_halfeye(
    data = ~subset(.x, drought_cat == "LS"),
    density = dens, .width = 0, alpha = 0.5,
    slab_color = "#E69F00", slab_linewidth = 1.1
  ) +
  ggdist::stat_halfeye(
    data = ~subset(.x, drought_cat == "MS"),
    density = dens, .width = 0, alpha = 0.5,
    slab_color = "#D55E00", slab_linewidth = 1.1
  ) +

  # Boxplots + means
  geom_boxplot(width = 0.35, linewidth = 0.8, color = "black", outlier.shape = NA) +
  stat_summary(fun = mean, geom = "point",
               shape = 21, size = 2, color = "black", fill = "white", stroke = 0.8) +

  # headroom for letters + prevent clipping after flip
  scale_y_continuous(
    breaks = seq(-10, 10, by = 2)#,
  ) +

  coord_flip(clip = "off") +
  theme_ddf_facet +

  # Align axis labels to nudged centers
  scale_x_continuous(
    breaks = c(1.0, 2.2, 3.4),
    labels = c("ND","LS","MS")
  ) +

  scale_fill_manual(values = c("ND"="#56B4E9","LS"="#E69F00","MS"="#D55E00")) +
  facet_wrap(~Veg, ncol = 2, nrow = 3) +

  labs(
    y = "Daily DDF (hours)", x = "",
  ) +

  # Tukey letters (aligned!)
  geom_text(
    data = letters_by_veg,
    inherit.aes = FALSE,
    aes(x = x_pos, y = y_lab, label = .label),
    fontface = "bold", size = 4
  )

print(VegSevCats)
ggsave(file.path(output_dir,"Figure_S4.png"),
       VegSevCats, width=10, height=8, dpi=300, bg="white", units="in")

```

