---
title: "Statistics"
author: "Kesondra Key"
date: "2026-03-29"
output: 
  pdf_document: default
---

IMPORTANT: 
Change "E:/Github/diurnal-divergence-factor/data/" to your own file location in the setup and on line 263

Purpose: Run statistics and generate tables
Note: Not all of the generated .csv's were used!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

#IMPORTANT: 
#Change "E:/Github/diurnal-divergence-factor/" to your own file location 
#input dir
daily_path <- "E:/Github/diurnal-divergence-factor/data/Masterfile.csv"
#output dir
output_dir <- "E:/Github/diurnal-divergence-factor/tables"


#Relevant packages installation
packages <- c(
  "dplyr", "tidyr", "ggplot2", "ggdist",
  "lubridate", "reshape2", "naniar", "BBmisc",
  "ggrepel", "gghighlight", "ggtext",
  "gridExtra", "grid","multcomp", "multcompView",
  "readr", "stringr", "tinytex", "knitr"
)

to_install <- packages[!packages %in% installed.packages()[, "Package"]]

if (length(to_install) > 0) {
  install.packages(to_install)
}

invisible(lapply(packages, library, character.only = TRUE))

# Libraries
library(dplyr); library(tidyr); library(ggplot2); library(ggdist)
library(lubridate); library(reshape2); library(naniar); library(BBmisc)
library(ggrepel); library(gghighlight); library(ggtext)
library(gridExtra); library(grid); library(multcomp); library(multcompView)
library(readr); library(stringr); library(tinytex); library(knitr)

Daily <- readr::read_csv(daily_path, show_col_types = FALSE)

Daily <- Daily %>%
  mutate(
    Date = as.Date(Date))

# Map p-values to stars
p_to_stars <- function(p) {
  if (is.na(p)) return("ns")
  if (p < 0.001) return("***")
  if (p < 0.01)  return("**")
  if (p < 0.05)  return("*")
  "ns"
}

# Tukey letters (for 3+ groups) per data.frame
tukey_letters <- function(dat, x, y) {
  dat <- dat[!is.na(dat[[x]]) & !is.na(dat[[y]]), , drop = FALSE]
  # group info
  g    <- factor(dat[[x]])
  levs <- levels(g)
  k    <- nlevels(g)

  make_tbl <- function(names_vec, labs_vec) {
    tibble::tibble(!!rlang::sym(x) := names_vec, .label = labs_vec)
  }

  if (k < 1) {
    return(make_tbl(character(0), character(0)))
  }
  if (k == 1) {
    return(make_tbl(levs, NA_character_))
  }
  if (k == 2) {
    p <- tryCatch(stats::t.test(dat[[y]] ~ g)$p.value, error = function(e) NA_real_)
    labs <- if (is.na(p) || p >= 0.05) c("a","a") else c("a","b")
    names(labs) <- levs
    return(make_tbl(names(labs), unname(labs)))
  }

  # k >= 3 → ANOVA + Tukey
  a  <- stats::aov(stats::as.formula(paste(y, "~", x)), data = dat)
  hs <- tryCatch(stats::TukeyHSD(a)[[1]][, "p adj"], error = function(e) NULL)

  if (is.null(hs)) {
    # fallback: give everyone "a"
    labs_full <- setNames(rep("a", k), levs)
  } else {
    tuk <- multcompView::multcompLetters(hs)$Letters   # named char vector
    # complete to all levels without modifyList()
    labs_full <- setNames(rep("a", k), levs)
    common <- intersect(names(tuk), names(labs_full))
    labs_full[common] <- tuk[common]
  }
  make_tbl(names(labs_full), unname(labs_full))
}

select <- dplyr::select
```

General information about the dataframe like number of sites and number of droughts per category
```{r General_Information, echo=FALSE}
periods <- Daily %>%
  filter(drought_cat %in% c("LS","MS"), !is.na(Cluster)) %>%
  distinct(sitename, Cluster, drought_cat, DroughtNo, ClusterStart, min_scaled_SWC)

periods %>%
  count(drought_cat, name = "n_periods")

periods %>%
  count(drought_cat, sitename, name = "n_periods") %>%
  group_by(drought_cat) %>%
  summarise(
    n_sites = n_distinct(sitename),
    median_periods_per_site = median(n_periods),
    max_periods_per_site = max(n_periods),
    .groups = "drop"
  )

# -----------------------------
# 1) One-line study summary numbers
# -----------------------------
n_sites <- Daily %>% distinct(sitename) %>% nrow()
n_sites
 Daily %>% nrow()
 summary(Daily)
 
n_site_years <- Daily %>% distinct(sitename, Year) %>% nrow()
n_site_years

# unique drought periods (clusters)
n_drought_periods <- Daily %>%
  filter(!is.na(Cluster)) %>%
  distinct(sitename, Cluster) %>%
  nrow()
n_drought_periods


cat(
  n_sites, "US flux tower sites were selected, covering",
  n_drought_periods, "unique drought periods across",
  n_site_years, "site-years.\n"
)

# -----------------------------
# 2) How many SITES per drought_cat / AI_cat / Veg
# (counts unique sites that appear in each category at least once)
# -----------------------------
sites_per_drought_cat <- Daily %>%
  filter(!is.na(drought_cat)) %>%
  distinct(sitename, drought_cat) %>%
  count(drought_cat, name = "n_sites") %>%
  arrange(desc(n_sites))
sites_per_drought_cat

periods_per_cat <- Daily %>%
  filter(drought_cat %in% c("LS","MS"), !is.na(Cluster)) %>%
  distinct(sitename, Cluster, drought_cat) %>%   # one row per drought period
  count(drought_cat, name = "n_periods") %>%
  arrange(desc(n_periods))
periods_per_cat


periods_per_cat <- Daily %>%
  filter(drought_cat %in% c("LS","MS"), !is.na(Cluster)) %>%
  distinct(sitename, Cluster, drought_cat) %>%   # one row per drought period
  count(drought_cat, name = "n_periods") %>%
  arrange(desc(n_periods))
periods_per_cat


sites_per_AI_cat <- Daily %>%
  filter(!is.na(AI_cat)) %>%
  distinct(sitename, AI_cat) %>%
  count(AI_cat, name = "n_sites") %>%
  arrange(desc(n_sites))
sites_per_AI_cat

sites_per_Veg <- Daily %>%
  filter(!is.na(Veg)) %>%
  distinct(sitename, Veg) %>%
  count(Veg, name = "n_sites") %>%
  arrange(desc(n_sites))
sites_per_Veg

# -----------------------------
# 3) "Separating drought_cat to the other categories"
# = how many UNIQUE sites per (AI_cat × drought_cat), (Veg × drought_cat)
# -----------------------------
sites_by_AIcat_droughtcat <- Daily %>%
  filter(!is.na(AI_cat), !is.na(drought_cat)) %>%
  distinct(sitename, AI_cat, drought_cat) %>%
  count(AI_cat, drought_cat, name = "n_sites") %>%
  arrange(AI_cat, drought_cat)
sites_by_AIcat_droughtcat


sites_by_Veg_droughtcat <- Daily %>%
  filter(!is.na(Veg), !is.na(drought_cat)) %>%
  distinct(sitename, Veg, drought_cat) %>%
  count(Veg, drought_cat, name = "n_sites") %>%
  arrange(Veg, drought_cat)
sites_by_Veg_droughtcat

```

\newpage
```{r table_1, echo=FALSE}
#This is the code to create Table 1
#IMPORTANT: update "E:/Github/diurnal-divergence-factor/" to your own filepath on line 263

#Table 1. 
#Summary of site-level metadata, DOI, and citation information for the datasets used in this study. 
#Abbreviations are as follows: lat, latitude; long, longitude; veg, vegetation; and AI, aridity index. 
#Mean aridity index methods are described in Section 2.3.1.

library(dplyr)
library(readr)

#daily_path <- "E:/Github/diurnal-divergence-factor/data/Masterfile.csv"
Daily <- readr::read_csv(daily_path, show_col_types = FALSE)

first_non_na <- function(x) {
  x <- x[!is.na(x)]
  if (length(x) == 0) NA else x[1]
}

tower_details <- Daily %>%
  dplyr::select(
    sitename,
    Latitude,
    Longitude,
    Veg,
    mean_AI,
    AI_cat,
    Terrain,
    data_source
  ) %>%
  distinct() %>%
  group_by(sitename) %>%
  dplyr::summarise(
    Latitude    = first_non_na(Latitude),
    Longitude   = first_non_na(Longitude),
    Veg         = first_non_na(Veg),
    mean_AI     = first_non_na(mean_AI),
    AI_cat      = first_non_na(AI_cat),
    Terrain     = first_non_na(Terrain),
    data_source = first_non_na(data_source),
    .groups = "drop"
  ) %>%
  arrange(sitename)

tower_details

readr::write_csv(tower_details, "E:/Github/diurnal-divergence-factor/tables/table_1.csv")
```


\newpage
# The Tables1 Section
This runs ANOVA, t-test, and description statistics, generate the information shown within the results, captions, and for tables 2, 3, 4, and S2:

# Table 2
## Descriptive statistics for DDF (hours) during drought and ND periods. Welch’s t-test: t(17109.79) = -9.02, p < 0.001; [mean] = 0.21 h, ~13 min
filename: desc_Drought_DDF.csv; ttest_DDF.csv

# Table 3
## Descriptive statistics for DDF (hours) for each drought severity category (ND, LS, MS)
filename: desc_drought_cat_DDF.csv 

# Table 4
## Descriptive statistics for DDF (hours) for each aridity category comparing each drought category (e.g., ND, drought)
filename: desc_AI_cat_DDF.csv 

# Table S2. 
## Summary statistical results for each vegetation (Veg) category
filename: desc_Veg_DDF.csv 

```{r Tables1, echo=FALSE, results="asis"}

y_var <- "DDF" 
Daily[[y_var]] <- suppressWarnings(as.numeric(Daily[[y_var]]))

# =========================
# 1) Helpers
# =========================
ttest_vs_ref_within <- function(dat, row_var, group_var, y_var, ref_level = "ND") {

  dat2 <- dat %>%
    dplyr::filter(!is.na(.data[[row_var]]), !is.na(.data[[group_var]]), !is.na(.data[[y_var]])) %>%
    dplyr::mutate(
      .row = as.factor(.data[[row_var]]),
      .grp = as.factor(.data[[group_var]]),
      .y   = as.numeric(.data[[y_var]])
    )

  out <- dat2 %>%
    dplyr::group_by(.row) %>%
    dplyr::group_modify(~{
      d <- .x

      # nothing to compare if ref is missing
      if (!(ref_level %in% levels(d$.grp))) {
        return(tibble::tibble())
      }

      comps <- setdiff(levels(d$.grp), ref_level)
      if (length(comps) == 0) return(tibble::tibble())

      res <- lapply(comps, function(cmp) {
        dd <- d %>% dplyr::filter(.grp %in% c(ref_level, cmp))

        x0 <- dd$.y[dd$.grp == ref_level]
        x1 <- dd$.y[dd$.grp == cmp]

        # require at least 2 obs per group for a t-test
        if (length(x0) < 2 || length(x1) < 2) {
          return(tibble::tibble(
            drought_cat = cmp,
            t = NA_real_, df = NA_real_, p = NA_real_,
            med_diff = stats::median(x1, na.rm = TRUE) - stats::median(x0, na.rm = TRUE)
          ))
        }

        tt <- tryCatch(stats::t.test(.y ~ .grp, data = dd), error = function(e) NULL)

        tibble::tibble(
          drought_cat = cmp,
          t  = if (!is.null(tt)) unname(tt$statistic) else NA_real_,
          df = if (!is.null(tt)) unname(tt$parameter) else NA_real_,
          p  = if (!is.null(tt)) unname(tt$p.value) else NA_real_,
          med_diff = stats::median(x1, na.rm = TRUE) - stats::median(x0, na.rm = TRUE)
        )
      })

      dplyr::bind_rows(res)
    }) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(
      p_adj = p,  # NO adjustment
      sig   = vapply(p, p_to_stars, character(1)),
      baseline = ref_level
    ) %>%
    dplyr::rename(!!row_var := .row)

  out
}


detect_site_var <- function(dat, candidates = c("sitename","site","SITE_ID","Site","site_id")) {
  hit <- candidates[candidates %in% names(dat)][1]
  if (length(hit) == 0 || is.na(hit)) stop("No site id column found. Add a site_var manually.")
  hit
}
site_var <- detect_site_var(Daily)

mean_ci_t <- function(x, conf = 0.95) {
  x <- x[is.finite(x)]
  n <- length(x)
  if (n < 2) return(c(NA_real_, NA_real_))
  m <- mean(x)
  s <- stats::sd(x)
  se <- s / sqrt(n)
  alpha <- 1 - conf
  tcrit <- stats::qt(1 - alpha/2, df = n - 1)
  c(m - tcrit * se, m + tcrit * se)
}

omega_sq_oneway <- function(aov_obj) {
  tab <- anova(aov_obj)
  ss_between <- tab[1, "Sum Sq"]
  df_between <- tab[1, "Df"]
  ss_within  <- tab[nrow(tab), "Sum Sq"]
  ms_within  <- tab[nrow(tab), "Mean Sq"]
  ss_total   <- ss_between + ss_within
  om <- (ss_between - df_between * ms_within) / (ss_total + ms_within)
  max(0, as.numeric(om))
}

summarise_by_group <- function(dat, group_var, y_var, site_var, conf = 0.95) {
  d <- dat %>%
    dplyr::filter(!is.na(.data[[group_var]]), !is.na(.data[[y_var]])) %>%
    dplyr::mutate(
      .grp = as.factor(.data[[group_var]]),
      .y   = as.numeric(.data[[y_var]])
    )
  
  out <- d %>%
  group_by(.grp) %>%
  summarise(
    sites_n = n_distinct(.data[[site_var]]),
    n_obs   = sum(is.finite(.y)),
    mean    = mean(.y, na.rm = TRUE),
    sd      = sd(.y, na.rm = TRUE),
    median  = median(.y, na.rm = TRUE),
    ci_low  = mean_ci_t(.y, conf)[1],
    ci_high = mean_ci_t(.y, conf)[2],
    .groups = "drop"
  ) %>%
  rename(Group = .grp)

 
  out
}

p_to_stars <- function(p) {
  if (is.na(p)) return("ns")
  if (p < 0.001) return("***")
  if (p < 0.01)  return("**")
  if (p < 0.05)  return("*")
  "ns"
}

run_ttest <- function(dat, group_var, y_var) {
  d <- dat %>%
    dplyr::filter(!is.na(.data[[group_var]]), !is.na(.data[[y_var]])) %>%
    dplyr::mutate(
      .grp = as.factor(.data[[group_var]]),
      .y   = as.numeric(.data[[y_var]])
    )

  levs <- levels(d$.grp)
  if (length(levs) != 2) return(NULL)

  x0 <- d$.y[d$.grp == levs[1]]
  x1 <- d$.y[d$.grp == levs[2]]

  tt <- tryCatch(stats::t.test(.y ~ .grp, data = d), error = function(e) NULL)

  tibble::tibble(
    test = "t-test",
    t  = if (!is.null(tt)) unname(tt$statistic) else NA_real_,
    df = if (!is.null(tt)) unname(tt$parameter) else NA_real_,
    p  = if (!is.null(tt)) unname(tt$p.value) else NA_real_,
    med_diff = median(x1, na.rm = TRUE) - median(x0, na.rm = TRUE),
    baseline = levs[1], comparison = levs[2]
  )
}

run_anova <- function(dat, group_var, y_var) {
  d <- dat %>%
    dplyr::filter(!is.na(.data[[group_var]]), !is.na(.data[[y_var]])) %>%
    dplyr::mutate(
      .grp = as.factor(.data[[group_var]]),
      .y   = as.numeric(.data[[y_var]])
    )

  if (nlevels(d$.grp) < 3) return(NULL)

  a <- stats::aov(.y ~ .grp, data = d)
  tab <- anova(a)

  tibble::tibble(
    test = "ANOVA",
    F   = as.numeric(tab[1, "F value"]),
    df1 = as.numeric(tab[1, "Df"]),
    df2 = as.numeric(tab[nrow(tab), "Df"]),
    p   = as.numeric(tab[1, "Pr(>F)"]),
    omega_sq = omega_sq_oneway(a)
  )
}

add_tukey_letters_to_desc <- function(dat, group_var, y_var, desc_tbl) {
  labs <- tukey_letters(dat, x = group_var, y = y_var)
  names(labs)[names(labs) == group_var] <- "Group"
  desc_tbl %>%
    dplyr::left_join(labs, by = "Group") %>%
    dplyr::rename(tukey = .label)
}

# =========================
# 2) What to run
# =========================
group_map <- tibble::tibble(
  group_var = c("Drought", "drought_cat", "Veg", "AI_cat"),
  Category  = c("Drought (Drought vs ND)", "Drought severity", "Vegetation type", "Aridity class")
) %>%
  dplyr::filter(group_var %in% names(Daily))

# =========================
# 3) Descriptive tables
# =========================
desc_tables <- list()

for (i in seq_len(nrow(group_map))) {
  gv <- group_map$group_var[i]
  catlab <- group_map$Category[i]

  desc <- summarise_by_group(Daily, gv, y_var, site_var, conf = 0.95) %>%
    dplyr::mutate(Category = catlab) %>%
    dplyr::select(Category, Group, sites_n, n_obs, mean, sd, median, ci_low, ci_high)

  k <- dplyr::n_distinct(Daily[[gv]], na.rm = TRUE)
  if (k >= 3) desc <- add_tukey_letters_to_desc(Daily, gv, y_var, desc)

  desc_tables[[gv]] <- desc
}

# =========================
# 4) Test tables (t-test or ANOVA)
# =========================
test_rows <- list()

for (i in seq_len(nrow(group_map))) {
  gv <- group_map$group_var[i]
  catlab <- group_map$Category[i]
  k <- dplyr::n_distinct(Daily[[gv]], na.rm = TRUE)

  if (k == 2) {
    test_rows[[gv]] <- run_ttest(Daily, gv, y_var) %>%
      dplyr::mutate(Category = catlab, group_var = gv)
  } else if (k >= 3) {
    test_rows[[gv]] <- run_anova(Daily, gv, y_var) %>%
      dplyr::mutate(Category = catlab, group_var = gv)
  }
}

tests <- dplyr::bind_rows(test_rows)

if (nrow(tests) > 0) {
tests <- tests %>%
  dplyr::mutate(sig = vapply(p, p_to_stars, character(1)))
}

ttest_table <- tests %>%
  dplyr::filter(test == "t-test") %>%
  dplyr::select(Category, group_var, test, t, df, p, sig, baseline, comparison, med_diff)
ttest_table

anova_table <- tests %>%
  dplyr::filter(test == "ANOVA") %>%
  dplyr::select(Category, group_var, test, F, df1, df2, p, sig, omega_sq)
anova_table

# =========================
# 5) Print tables
# =========================
for (nm in names(desc_tables)) {
  cap <- paste0("Descriptive stats by ", nm, " (response = ", y_var, "). CI = 95% t-based CI of the mean.")
  print(knitr::kable(desc_tables[[nm]], digits = 3, caption = cap, booktabs = TRUE))
  cat("\n\n")
}

if (nrow(ttest_table) > 0) {
  print(knitr::kable(ttest_table, digits = 4, caption = "t-tests", booktabs = TRUE))
  cat("\n\n")
}

if (nrow(anova_table) > 0) {
  print(knitr::kable(anova_table, digits = 4, caption = "ANOVAs + omega squared", booktabs = TRUE))
  cat("\n\n")
}


# =========================
# 6) Write CSVs
# =========================
# Descriptives
for (nm in names(desc_tables)) {
  readr::write_csv(desc_tables[[nm]], file.path(output_dir, paste0("desc_", nm, "_", y_var, ".csv")))
}
# Tests
if (nrow(ttest_table) > 0) readr::write_csv(ttest_table, file.path(output_dir, paste0("ttest_", y_var, ".csv")))
if (nrow(anova_table) > 0) readr::write_csv(anova_table, file.path(output_dir, paste0("anova_", y_var, ".csv")))
```

\newpage
# The Tables2 Section
Drought severity - this runs ANOVA, t-test, and description statistics, generate the information shown within the results, captions, and for tables 4, 5, S3, and S4:


# Table 4. 
## Descriptive statistics for DDF (hours) for each aridity category comparing each drought category (e.g., ND, drought)
filename: desc_AI_cat_x_drought_DDF.csv

# Table 5. 
## Descriptive statistics for DDF (hours) for each aridity category comparing each drought severity category
filename: desc_AI_cat_x_drought_cat_DDF.csv

# Table S3. 
## Summary statistic results for each vegetation (Veg) type under non-drought (ND) and drought conditions
filename: desc_Veg_x_drought_DDF.csv

# Table S4. 
## Summary statistic results for each vegetation (Veg) type under non-drought (ND), less severe (LS) drought, and more severe (MS) drought conditions
filename: desc_Veg_x_drought_cat_DDF.csv

```{r Tables2, echo=FALSE, results="asis"}
y_var <- "DDF"  

Daily[[y_var]] <- suppressWarnings(as.numeric(Daily[[y_var]]))

if ("Drought" %in% names(Daily)) {
  Daily <- Daily %>% mutate(Drought = factor(Drought, levels = c("ND","Drought")))
}

# -------------------------
# HELPERS
# -------------------------
run_tukey_within <- function(dat, row_var, group_var, y_var) {
  dat %>%
    dplyr::filter(!is.na(.data[[row_var]]), !is.na(.data[[group_var]]), !is.na(.data[[y_var]])) %>%
    dplyr::mutate(
      .row = as.factor(.data[[row_var]]),
      .grp = as.factor(.data[[group_var]]),
      .y   = as.numeric(.data[[y_var]])
    ) %>%
    dplyr::group_by(.row) %>%
    dplyr::group_modify(~{
      d <- .x

      if (nlevels(d$.grp) < 3) {
        return(tibble::tibble())
      }

      fit <- stats::aov(.y ~ .grp, data = d)
      tk  <- stats::TukeyHSD(fit, which = ".grp")

      as.data.frame(tk$.grp) %>%
        tibble::rownames_to_column("comparison") %>%
        tibble::as_tibble() %>%
        dplyr::rename(p_adj = `p adj`) %>%
        tidyr::separate(comparison, into = c("group1", "group2"), sep = "-", remove = FALSE)
    }) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(
      sig = vapply(p_adj, p_to_stars, character(1))
    ) %>%
    dplyr::rename(!!row_var := .row)
}

detect_site_var <- function(dat, candidates = c("sitename","site","SITE_ID","Site","site_id")) {
  hit <- candidates[candidates %in% names(dat)][1]
  if (length(hit) == 0 || is.na(hit)) stop("No site id column found. Add a site_var manually.")
  hit
}
site_var <- detect_site_var(Daily)

mean_ci_t <- function(x, conf = 0.95) {
  x <- x[is.finite(x)]
  n <- length(x)
  if (n < 2) return(c(NA_real_, NA_real_))
  m <- mean(x); s <- sd(x); se <- s / sqrt(n)
  tcrit <- qt(1 - (1-conf)/2, df = n - 1)
  c(m - tcrit * se, m + tcrit * se)
}

omega_sq_oneway <- function(aov_obj) {
  tab <- anova(aov_obj)
  ss_between <- tab[1, "Sum Sq"]
  df_between <- tab[1, "Df"]
  ss_within  <- tab[nrow(tab), "Sum Sq"]
  ms_within  <- tab[nrow(tab), "Mean Sq"]
  ss_total   <- ss_between + ss_within
  om <- (ss_between - df_between * ms_within) / (ss_total + ms_within)
  max(0, as.numeric(om))
}

p_to_stars <- function(p) {
  if (is.na(p)) return("ns")
  if (p < 0.001) return("***")
  if (p < 0.01)  return("**")
  if (p < 0.05)  return("*")
  "ns"
}

summarise_two_way <- function(dat, row_var, col_var, y_var, site_var, conf = 0.95) {
  dat %>%
    filter(!is.na(.data[[row_var]]), !is.na(.data[[col_var]]), !is.na(.data[[y_var]])) %>%
    mutate(
      .row = as.factor(.data[[row_var]]),
      .col = as.factor(.data[[col_var]]),
      .y   = as.numeric(.data[[y_var]])
    ) %>%
    group_by(.row, .col) %>%
    summarise(
      sites_n = n_distinct(.data[[site_var]]),
      n_obs   = sum(is.finite(.y)),
      mean    = mean(.y, na.rm = TRUE),
      sd      = sd(.y, na.rm = TRUE),
      median  = median(.y, na.rm = TRUE),
      ci_low  = mean_ci_t(.y, conf)[1],
      ci_high = mean_ci_t(.y, conf)[2],
      .groups = "drop"
    ) %>%
    rename(!!row_var := .row, !!col_var := .col)
}

# run t-test within each stratum of row_var: compare drought (2 groups)
ttest_within <- function(dat, row_var, group_var, y_var) {
  dat %>%
    filter(!is.na(.data[[row_var]]), !is.na(.data[[group_var]]), !is.na(.data[[y_var]])) %>%
    mutate(
      .row = as.factor(.data[[row_var]]),
      .grp = as.factor(.data[[group_var]]),
      .y   = as.numeric(.data[[y_var]])
    ) %>%
    group_by(.row) %>%
    group_modify(~{
      d <- .x
      levs <- levels(d$.grp)
      if (length(levs) != 2) {
        return(tibble(t=NA_real_, df=NA_real_, p=NA_real_, med_diff=NA_real_, baseline=NA_character_, comparison=NA_character_))
      }
      x0 <- d$.y[d$.grp == levs[1]]
      x1 <- d$.y[d$.grp == levs[2]]
      if (length(x0) < 2 || length(x1) < 2) {
        return(tibble(t=NA_real_, df=NA_real_, p=NA_real_,
                      med_diff = median(x1, na.rm=TRUE) - median(x0, na.rm=TRUE),
                      baseline=levs[1], comparison=levs[2]))
      }
      tt <- tryCatch(t.test(.y ~ .grp, data = d), error = function(e) NULL)
      tibble(
        t  = if (!is.null(tt)) unname(tt$statistic) else NA_real_,
        df = if (!is.null(tt)) unname(tt$parameter) else NA_real_,
        p  = if (!is.null(tt)) unname(tt$p.value) else NA_real_,
        med_diff = median(x1, na.rm=TRUE) - median(x0, na.rm=TRUE),
        baseline = levs[1], comparison = levs[2]
      )
    }) %>%
    ungroup() %>%
    dplyr::mutate(
      p_adj = p,                 
      sig   = vapply(p, p_to_stars, character(1))
    ) %>%
    dplyr::rename(!!row_var := .row)

}

# run one-way ANOVA within each stratum of row_var on group_var (3+ groups)
anova_within <- function(dat, row_var, group_var, y_var) {
  dat %>%
    filter(!is.na(.data[[row_var]]), !is.na(.data[[group_var]]), !is.na(.data[[y_var]])) %>%
    mutate(
      .row = as.factor(.data[[row_var]]),
      .grp = as.factor(.data[[group_var]]),
      .y   = as.numeric(.data[[y_var]])
    ) %>%
    group_by(.row) %>%
    group_modify(~{
      d <- .x
      if (nlevels(d$.grp) < 3) {
        return(tibble(F=NA_real_, df1=NA_real_, df2=NA_real_, p=NA_real_, omega_sq=NA_real_))
      }
      a <- aov(.y ~ .grp, data = d)
      tab <- anova(a)
      tibble(
        F   = as.numeric(tab[1, "F value"]),
        df1 = as.numeric(tab[1, "Df"]),
        df2 = as.numeric(tab[nrow(tab), "Df"]),
        p   = as.numeric(tab[1, "Pr(>F)"]),
        omega_sq = omega_sq_oneway(a)
      )
    }) %>%
    ungroup() %>%
    dplyr::mutate(
      p_adj = p,                        # no adjustment
      sig   = vapply(p, p_to_stars, character(1))
    ) %>%
    dplyr::rename(!!row_var := .row)

}

# -------------------------
# 1) VEG × drought (ND vs Drought) + t-tests within Veg
# -------------------------
if (all(c("Veg","Drought") %in% names(Daily))) {
  desc_veg_drought <- summarise_two_way(Daily, "Veg", "Drought", y_var, site_var) %>%
    dplyr::mutate(Category = "Veg") %>%
    dplyr::select(Category, Veg, Drought, sites_n, n_obs, mean, sd, median, ci_low, ci_high)
  
  tests_veg_drought <- ttest_within(Daily, "Veg", "Drought", y_var) %>%
    dplyr::mutate(Category = "Veg") %>%
    dplyr::select(Category, Veg, t, df, p, p_adj, sig, baseline, comparison, med_diff)
  
  desc_veg_droughtcat <- summarise_two_way(Daily, "Veg", "drought_cat", y_var, site_var) %>%
    dplyr::mutate(Category = "Veg") %>%
    dplyr::select(Category, Veg, drought_cat, sites_n, n_obs, mean, sd, median, ci_low, ci_high)
  
  tests_veg_droughtcat <- anova_within(Daily, "Veg", "drought_cat", y_var) %>%
    dplyr::mutate(Category = "Veg") %>%
    dplyr::select(Category, Veg, F, df1, df2, p, p_adj, sig, omega_sq)
  
  print(knitr::kable(desc_veg_drought, digits = 3, booktabs = TRUE,
                     caption = paste0("Veg × drought (response=", y_var, "); CI = 95% t-based CI of mean")))
  cat("\n\n")
  print(knitr::kable(tests_veg_drought, digits = 4, booktabs = TRUE,
                     caption = "t-tests (within each Veg): ND vs Drought (no p adjustment)"))
  cat("\n\n")

  readr::write_csv(desc_veg_drought, file.path(output_dir, paste0("desc_Veg_x_drought_", y_var, ".csv")))
  readr::write_csv(tests_veg_drought, file.path(output_dir, paste0("ttest_within_Veg_drought_", y_var, ".csv")))
}

# -------------------------
# 2) VEG × drought_cat (ND/LS/MS...) + ANOVAs within Veg
# -------------------------
if (all(c("Veg","drought_cat") %in% names(Daily))) {
  desc_veg_droughtcat <- summarise_two_way(Daily, "Veg", "drought_cat", y_var, site_var) %>%
    mutate(Category = "Veg") %>%
    dplyr::select(Category, Veg, drought_cat, sites_n, n_obs, mean, sd, median, ci_low, ci_high)

  tests_veg_droughtcat <- anova_within(Daily, "Veg", "drought_cat", y_var) %>%
    mutate(Category = "Veg") %>%
    select(Category, Veg, F, df1, df2, p, sig, omega_sq)

  print(knitr::kable(desc_veg_droughtcat, digits = 3, booktabs = TRUE,
                     caption = paste0("Veg × drought_cat (response=", y_var, "); CI = 95% t-based CI of mean")))
  cat("\n\n")
  print(knitr::kable(tests_veg_droughtcat, digits = 4, booktabs = TRUE,
                     caption = "ANOVAs (within each Veg): drought_cat (no p adjustment)"))
  cat("\n\n")

  readr::write_csv(desc_veg_droughtcat, file.path(output_dir, paste0("desc_Veg_x_drought_cat_", y_var, ".csv")))
  readr::write_csv(tests_veg_droughtcat, file.path(output_dir, paste0("anova_within_Veg_drought_cat_", y_var, ".csv")))
}


if (all(c("Veg", "drought_cat") %in% names(Daily))) {

  tests_veg_droughtcat_pairwise <- ttest_vs_ref_within(
    Daily, row_var = "Veg", group_var = "drought_cat", y_var = y_var, ref_level = "ND"
  ) %>%
    dplyr::mutate(Category = "Veg") %>%
    dplyr::select(Category, Veg, drought_cat, t, df, p, p_adj, sig, baseline, med_diff)

  print(knitr::kable(
    tests_veg_droughtcat_pairwise, digits = 4, booktabs = TRUE,
    caption = "Pairwise t-tests within each Veg: drought_cat vs ND (no p adjustment)"
  ))
  cat("\n\n")

  readr::write_csv(
    tests_veg_droughtcat_pairwise,
    file.path(output_dir, paste0("ttest_within_Veg_drought_cat_vs_ND_", y_var, ".csv"))
  )
}

if (all(c("AI_cat", "drought_cat") %in% names(Daily))) {

  tests_ai_droughtcat_pairwise <- ttest_vs_ref_within(
    Daily, row_var = "AI_cat", group_var = "drought_cat", y_var = y_var, ref_level = "ND"
  ) %>%
    dplyr::mutate(Category = "AI_cat") %>%
    dplyr::select(Category, AI_cat, drought_cat, t, df, p, p_adj, sig, baseline, med_diff)

  print(knitr::kable(
    tests_ai_droughtcat_pairwise, digits = 4, booktabs = TRUE,
    caption = "Pairwise t-tests within each AI_cat: drought_cat vs ND (no p adjustment)"
  ))
  cat("\n\n")

  readr::write_csv(
    tests_ai_droughtcat_pairwise,
    file.path(output_dir, paste0("ttest_within_AI_cat_drought_cat_vs_ND_", y_var, ".csv"))
  )
}

if (all(c("Veg", "drought_cat") %in% names(Daily))) {

  Daily <- Daily %>%
    dplyr::mutate(
      Veg = factor(Veg),
      drought_cat = factor(drought_cat, levels = c("ND", "LS", "MS"))
    )

  tests_veg_droughtcat_tukey <- run_tukey_within(
    Daily,
    row_var = "Veg",
    group_var = "drought_cat",
    y_var = y_var
  ) %>%
    dplyr::mutate(Category = "Veg") %>%
    dplyr::select(Category, Veg, comparison, group1, group2, diff, lwr, upr, p_adj, sig)

  print(knitr::kable(
    tests_veg_droughtcat_tukey,
    digits = 4,
    booktabs = TRUE,
    caption = "Tukey HSD pairwise comparisons within each Veg: drought_cat"
  ))
  cat("\n\n")

  readr::write_csv(
    tests_veg_droughtcat_tukey,
    file.path(output_dir, paste0("tukey_within_Veg_drought_cat_", y_var, ".csv"))
  )
}

if (all(c("AI_cat", "drought_cat") %in% names(Daily))) {

  Daily <- Daily %>%
    dplyr::mutate(
      AI_cat = factor(AI_cat),
      drought_cat = factor(drought_cat, levels = c("ND", "LS", "MS"))
    )

  tests_ai_droughtcat_tukey <- run_tukey_within(
    Daily,
    row_var = "AI_cat",
    group_var = "drought_cat",
    y_var = y_var
  ) %>%
    dplyr::mutate(Category = "AI_cat") %>%
    dplyr::select(Category, AI_cat, comparison, group1, group2, diff, lwr, upr, p_adj, sig)

  print(knitr::kable(
    tests_ai_droughtcat_tukey,
    digits = 4,
    booktabs = TRUE,
    caption = "Tukey HSD pairwise comparisons within each AI_cat: drought_cat"
  ))
  cat("\n\n")

  readr::write_csv(
    tests_ai_droughtcat_tukey,
    file.path(output_dir, paste0("tukey_within_AI_cat_drought_cat_", y_var, ".csv"))
  )
}

# -------------------------
# Repeat for AI_cat × drought / AI_cat × drought_cat
# -------------------------
if (all(c("AI_cat","Drought") %in% names(Daily))) {
  desc_ai_drought <- summarise_two_way(Daily, "AI_cat", "Drought", y_var, site_var) %>%
    mutate(Category = "AI_cat") %>%
    select(Category, AI_cat, Drought, sites_n, n_obs, mean, sd, median, ci_low, ci_high)

  tests_ai_drought <- ttest_within(Daily, "AI_cat", "Drought", y_var) %>%
    mutate(Category = "AI_cat") %>%
    select(Category, AI_cat, t, df, p, sig, baseline, comparison, med_diff)

  print(knitr::kable(desc_ai_drought, digits = 3, booktabs = TRUE,
                     caption = paste0("AI_cat × drought (response=", y_var, "); CI = 95% t-based CI of mean")))
  cat("\n\n")
  print(knitr::kable(tests_ai_drought, digits = 4, booktabs = TRUE,
                     caption = "t-tests (within each AI_cat): ND vs Drought (no p adjustment)"))
  cat("\n\n")

  readr::write_csv(desc_ai_drought, file.path(output_dir, paste0("desc_AI_cat_x_drought_", y_var, ".csv")))
  readr::write_csv(tests_ai_drought, file.path(output_dir, paste0("ttest_within_AI_cat_drought_", y_var, ".csv")))
}

if (all(c("AI_cat","drought_cat") %in% names(Daily))) {
  desc_ai_droughtcat <- summarise_two_way(Daily, "AI_cat", "drought_cat", y_var, site_var) %>%
    mutate(Category = "AI_cat") %>%
    select(Category, AI_cat, drought_cat, sites_n, n_obs, mean, sd, median, ci_low, ci_high)

  tests_ai_droughtcat <- anova_within(Daily, "AI_cat", "drought_cat", y_var) %>%
    mutate(Category = "AI_cat") %>%
    select(Category, AI_cat, F, df1, df2, p, sig, omega_sq)

  print(knitr::kable(desc_ai_droughtcat, digits = 3, booktabs = TRUE,
                     caption = paste0("AI_cat × drought_cat (response=", y_var, "); CI = 95% t-based CI of mean")))
  cat("\n\n")
  print(knitr::kable(tests_ai_droughtcat, digits = 4, booktabs = TRUE,
                     caption = "ANOVAs (within each AI_cat): drought_cat (no p adjustment)"))
  cat("\n\n")

  readr::write_csv(desc_ai_droughtcat, file.path(output_dir, paste0("desc_AI_cat_x_drought_cat_", y_var, ".csv")))
  readr::write_csv(tests_ai_droughtcat, file.path(output_dir, paste0("anova_within_AI_cat_drought_cat_", y_var, ".csv")))
}
```

